---
title: "Intro to Keras"
author: "Elizabeth Karan"
date: "2/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All of the information here is from https://keras.rstudio.com

## What is Keras?
Keras is a high-level neural networks API used for Deep Learning. Deep learning is machine learning with *layers*. Keras can run Tensorflow, CNTK, Theano and other backends. **It uses the Tensorflow backend engine as a default.** It also works with R, which is super nice and convenient. 


## Installation Intructions (DO THIS BEFORE YOU RUN ANYTHING)

**Make sure you already have devtools installed**  
library(devtools)

To install Keras you need to run the following:  

devtools::install_github("rstudio/keras")  

Then run this:  

library(keras)  
install_keras()  


## MNIST Tutorial

**Prepare the data**

Learning to recognize handwritten digits with the MNIST dataset. The dataset contains x and y: x is a 3-d array (images,width,height) of grayscale values and y is an integer vector with values ranging from 0 to 9 that represent labels for each category (in this case each represents which type of digit, 1 through 9, is written). 
```{r prep}
library(keras)
## the MNIST dataset is included with Keras
mnist <- dataset_mnist()

## set up the test-train split
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y
```

Reshape into matrix with width and height in a single dimension; each 28x28 image is flattened into a vector with a length of 784. Rescale: the grayscale values are converted from integers ranging between 0 to 255 into floating point values ranging between 0 and 1. 
```{r reshapeandrescale}
## reshape 
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
## rescale
x_train <- x_train / 255
x_test <- x_test / 255
```

Convert the vectors y vectors into binary class matricies
```{r yvect}
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```


**Define the model**

Create a sequential model (a linear stack of layers) and add layers using '>%>'
```{r model}
model <- keras_model_sequential() 
model %>% 
  # shape of the input data is a numeric vector representing a grayscale image with length of 784 
  layer_dense(units = 256, activation = 'relu', input_shape = c(784)) %>% 
  layer_dropout(rate = 0.4) %>% 
  layer_dense(units = 128, activation = 'relu') %>%
  layer_dropout(rate = 0.3) %>%
  # outputs a length 10 numeric vector w/ probabilities for each digit
  layer_dense(units = 10, activation = 'softmax')

summary(model)
```

Compile the model
```{r compile}
model %>% compile(
  # loss function
  loss = 'categorical_crossentropy',
  # optimizer
  optimizer = optimizer_rmsprop(),
  # metrics
  metrics = c('accuracy')
)
```


**Training and Evaluation**

Use the fit() function to train the model for 30 epochs using batches of 128 images. This takes a few minutes to run. 
```{r train}
history <- model %>% fit(
  x_train, y_train, 
  epochs = 30, batch_size = 128, 
  validation_split = 0.2
)
```

Plot history, which includes loss and accuracy metrics
```{r plot}
plot(history)
```

Evaluate the model's performance, in terms of loss and accuracy, using the test data
```{r performance}
model %>% evaluate(x_test, y_test)
```

Make predictions on the new data
```{r predict}
model %>% predict_classes(x_test)
```

